---
title: "Optimization of real state investment portfolio using R"
bibliography: EVancouver.bib
csl: vancouver.csl
author: "Ariel Levy$^{1}$; Marcus Antonio Cardoso Ramalho$^{1}$"
toc: TRUE
toc-location: left 
toc-depth: 2
format: 
  revealjs:
    theme: night
    chalkboard: true
    incremental: true
    slide-number: true
    show-slide-number: print
    preview-links: auto
    logo: xornada.png
    footer: <https://www.ppgad.uff.br/>
date: '`r Sys.Date()`'
---


## Introduction {.smaller}

::: {#tabset1 .panel-tabset}
## Context

FII - is kind of real estate investment in Brazil

-   Monthly 95% dividend payments

-   Growth by public offers

-   Several categories

## Popularity



```{r}
#| echo: false
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
suppressMessages(library(readxl))
suppressMessages(library(highcharter))
suppressMessages(library(tidyquant))
suppressMessages(library(tibbletime))
suppressMessages(library(quantmod))
suppressMessages(library(PerformanceAnalytics))
suppressMessages(library(scales))
library(timetk)
library(readxl)
library(here)
```

::: {#fig-elephants layout-ncol="2"}

```{r}
data_inv <-seq(as.Date("2009-12-31"), as.Date("2023-08-31"), by="years")#cria o vetor de datas

investors <-as.numeric(c("12000","21000",
                         "36000","102000",
                         "101000","93000",
                         "89000","89000",
                         "121000","208000",
                         "645000","1172000",
                         "1546000","1837000"))

n_investors<-data.frame(data_inv, investors) %>% 
  mutate(t_investors=investors/1000)
ggplot(n_investors) +
  aes(x = data_inv, y = t_investors) +
  geom_col(fill = "deepskyblue",position = ) +
  geom_text(aes(label=t_investors),vjust=-0.2)+
  labs(title = "Number of investors",
       subtitle = "Evolution of the number of investors - custody position (Thousands)" ,
       x="Year", y="Number of investor")+
       ggthemes::theme_few()
```


```{r}
#dataframe evolução anual dos FIIs

date_fiis <-seq(as.Date("2017-12-31"), as.Date("2023-07-31"), by="years")

Listed<-as.numeric(c("156","190",
                     "250","311",
                     "402","431"))

Total<-as.numeric(c("333","381",
                    "464","555",
                    "697","756"))

anual_fiis_ev<-data.frame(date_fiis,Listed,Total)#constroi o dataframe

#deixa o dataframe na forma longa para apresentar os graficos de barra na forma empilhada
an_fii_ev_plot<-pivot_longer(anual_fiis_ev,-date_fiis) 

#Number of funds and primary market
 g2<- ggplot(an_fii_ev_plot) +
  aes(x = date_fiis, fill = name, y = value) +
  geom_col(position = "dodge",stat='identity') +
  ggfittext::geom_bar_text(position="dodge")+
  scale_fill_manual(values = c(Listed= "lightblue",
    Total = "deepskyblue")) +
    labs(title = "Number of funds and primary market",
    subtitle = "Yearly evolution of the number of FIIs" ,
    x="Year", y="Total de fundos CVM / Total Funds CVM",fill="")+
  ggthemes::theme_few()
  
 g2
```
:::

## The problem

-   Maximize portfolio returns

-   Verify interest rate impact

## Choices

-   liquidity

-   categories we left out

## 



```{r}

comp_ifix_selic <- readxl::read_excel("dados/comp_ifix_selic.xlsx",
    col_types = c("date", "text", "numeric"))


ggplot(comp_ifix_selic) +
  aes(x = date, y = returns, colour = asset) +
  geom_line(size = 0.5) +
  scale_color_hue(direction = 1) +
  theme_minimal()+
  labs(title = "Comparative returns IFIX vs SELIC - 2013-2022")
```
:::

## Metodolgia

::: {#tabset2 .panel-tabset}
## Data

::: {#fig-elephants layout-ncol="2"}




Famous Elephants
:::

## Packages

## Recepies
:::

## Discussão

tabset3 .panel-tabset}

## Theory

## Hipothesys

## Ongoing Research

## Conclusion

-   Interst rates impact
-   Sharpe Ratio
-   Limitations \_ Future studies



